---
title: "Project-12"
format: html
---

```{r, message=FALSE, warning=FALSE}
here::i_am("DM-12.Rproj")
library(here)
library(vroom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(patchwork)
library(stringr)
```

## Data loading 
```{r, warning=FALSE, message=FALSE}
educ_expenditure <- vroom(here("data", "Expenditure on education.csv"), 
                          locale=locale(decimal_mark = "."))
GDP_capita <- vroom(here("data", "GDP per capita.csv"), 
                    locale=locale(decimal_mark = "."))
life_expectancy <- vroom(here("data", "life_expectancy.csv"), 
                         locale=locale(decimal_mark = "."))
PressFreedom <- vroom(here("data", "PressFreedom.csv"), 
                      locale=locale(decimal_mark = "."))
GDP <- vroom(here("data", "GDP.csv"), 
             locale=locale(decimal_mark = "."))
population <- vroom(here("data", "population.csv"), 
                    locale=locale(decimal_mark = "."))
enrolment  <- vroom(here("data", "enrolment.csv"), 
                    locale=locale(decimal_mark = "."))
illiteracy <- vroom(here("data", "illiteracy.csv"), 
                    locale=locale(decimal_mark = "."))
electricity <- vroom(here("data", "Electricity.csv"), 
                     locale=locale(decimal_mark = "."), 
                     na="..")
religion <- vroom(here("data", "Religion.csv"), 
                  locale=locale(decimal_mark = "."))
institution <- vroom(here("data", "stability-democratic-institutions-index.csv"), 
                     locale=locale(decimal_mark = "."))
rural <- vroom("data/Rural.csv", col_types=cols(
   DEMO_IND = col_character(),
   Indicator = col_character(),
   LOCATION = col_character(),
   Country = col_character(),
   TIME = col_double(),
   Time = col_double(),
   Value = col_double(),
   `Flag Codes` = col_character(),
   Flags = col_character(),
   .delim = ","))
```

## Github 

[Here](https://github.com/GraceBeyoko/DM-12.git) is the link to our Github repository.

## Sources

We used the table from the Our World in Data website for the Stability of democratic institutions index variable, available [here](https://ourworldindata.org/grapher/stability-democratic-institutions-index-bti?tab=table).

The Press freedom index and the Access to electricity tables are both sourced from the World Bank website, respectively available [here](https://govdata360.worldbank.org/indicators/h3f86901f?country=BRA&indicator=32416&viz=line_chart&years=2002,2023) and [here](https://databank.worldbank.org/reports.aspx?source=2&series=EG.ELC.ACCS.ZS&country=#).

The table for the religion per country comes from the the UNdata website, available [here](https://data.un.org/Data.aspx?d=POP&f=tableCode%3A28). 

The rest of our data (GDP, GDP per capita, rural population, life expectancy at birth, public expenditure allocated towards education, the total net enrolment rate by level of education and illiterate population) comes from the UNESCO website, available [here](http://data.uis.unesco.org/?lang=en&SubSessionId=4db844f6-01dd-458d-84db-7a36d4601fa2&themetreeid=-200). To find each of them in the website, we did the following :

- for GDP and GDP per capita, we selected the "demographic and socio-economic" category and then the "socio-economic" section. 

- for rural population and life expectancy at birth, we selected the "demographic and socio-economic" category and then the "demographic" section. 

- For expenditure allocated towards education, we selected the "Education" category then the "Other policy relevant indicators" section and then the "Government expenditure on education as a percentage of GDP" sub-section

- For the total net enrolment rate by level of education, we selected the "Education" category then the "Other policy relevant indicators" section and then "Number of students and enrolment/attendance rates by level of education" and then the "Enrolment ratios" sub-section. Here, we choose each level of education but only for both sexes. 

- For illiterate population, we selected the "Education" category then the "Other policy relevant indicators" section then the "Number of illiterates/percentage of illiterate who are female" sub-section and then we select "illiterate population".

## Sources description

Our World in Data is a free online publication founded by Max Roser, an economist, and directed by the University of Oxford. They focus on providing comprehensive database and data visualizations to make information about global development more accessible. 

The World Bank is a financial institution that provides loans and grants to the governments of developing countries for the purpose of pursuing investments projects. On top of that, they collect, analyse, and publish a wide range of data on various aspects of the economies and societies of their member countries. These data are made available to the public and are widely used by researchers, policy-makers, businesses, and the general public.

UNdata is an search engine that provides access to a wide range of statistical databases and other data sources covering various economic, social, demographic, and environmental topics. This service has been developed in partnership with the United Nations Statistics Division and the Statistics Sweden and the Swedish International Development Cooperation Agency in 2008.

The United Nations Educational, Scientific and Cultural Organization (UNESCO) is a specialized agency of the United Nations established in 1945. The organization mission is to contribute to the building of peace, the eradication of poverty, sustainable development and intercultural dialogue through education, sciences and culture. One of their objectives is to ease access to data. That is why, they also produce database about education, culture, science, and communication.

## Data Cleaning 

We are now cleaning the variables from the UNESCO website.


All the datasets from the UNESCO site have the same structure, as seen below:

The columns' names of the Expenditure on Education dataset are : `r educ_expenditure |> names()`.

The columns' names of the GDP per capita dataset are :`r GDP_capita |> names()`.

The columns' names of the GDP dataset are :`r GDP |> names()`.

The columns' names of the Population dataset are :`r population |> names()`.

The columns' names of the Enrolment by level of Education dataset are :`r enrolment |> names()`.

The columns' names of the Illiteracy dataset are :`r illiteracy |> names()`.

The columns' names of the Rural population share dataset are :`r rural |> names()`.

That is why we can apply the same function to each to the UNESCO datasets. Here we choose to eliminate Flags, Flag Codes and the indicator codes because they do not fall in our research scope. On top of that, we can see that there are two variables, Times and TIMES, that are duplicate of each other as seen in the following. So we will also removes the TIMES column.

```{r}
sum(educ_expenditure$Time == educ_expenditure$TIME) == nrow(educ_expenditure)
```

Here, we see that this column contains the same value each time. So, we will also remove this column.
```{r}
educ_expenditure |>
  distinct(NATMON_IND) |> 
  knitr::kable()
```

We will now apply the following function to all of the UNESCO data. 
```{r}
datacleaning_unesco <- function(data){
    data |> select(-`Flag Codes`, - Flags, - TIME, -ends_with("IND"))
}
```

```{r}
educ_expenditure <- datacleaning_unesco(educ_expenditure)
GDP_capita <- datacleaning_unesco(GDP_capita)
GDP <- datacleaning_unesco(GDP)
population <- datacleaning_unesco(population)
enrolment  <- datacleaning_unesco(enrolment)
illiteracy <-datacleaning_unesco(illiteracy)
rural <- datacleaning_unesco(rural)
```





We now are doing the data cleaning of the electricity variable. 
```{r}
electricity |> 
  names()|>
  knitr::kable()
```

Here, we want to see what values these columns take. 
```{r}
electricity |> 
  distinct(`Series Code`, `Series Name`) |> 
  knitr::kable()
```

Because access to electricity is our variable of interest, we filter the dataset.
```{r}
electricity <- electricity |>
  filter(`Series Name`== "Access to electricity (% of population)")|> 
  select(- `Series Code`)
```

We now are going to rename the columns for them to match the other variables from the UNESCO.
```{r}
electricity <- electricity |> 
  rename("Country" = "Country Name", "LOCATION"= "Country Code", "Indicator"="Series Name")
```

We do pivot_longer in order to have all of our datasets in the same format.
```{r}
electricity <- electricity |> 
  pivot_longer(!(Indicator| Country| LOCATION)) |> 
  rename("Time"= "name", "Value"= "value") |> 
  mutate("Time"= as.numeric(substr(Time, 1, 4)))
```





We now are cleaning the PressFreedom variable.
```{r}
PressFreedom |> 
  names() |>
  knitr::kable()
```

```{r}
PressFreedom |> 
  distinct(`Subindicator Type`, `Indicator ID`) |>
  knitr::kable()
```

```{r}
PressFreedom <- PressFreedom |> select(- `Indicator ID`, - `Subindicator Type`)
```

```{r}
PressFreedom <- PressFreedom |> rename("Country" = "Country Name", "LOCATION"= "Country ISO3")
```

```{r}
PressFreedom |> 
  count(Indicator) |> 
  knitr::kable()
```

Because our PressFreedom table contains two indicators we split the table in order to have two tables representing each indicator. 
```{r}
PressFreedomIndex <- PressFreedom |> filter(Indicator == "Press Freedom Index")
PressFreedomRank <- PressFreedom |> filter(Indicator == "Press Freedom Rank")
```

```{r}
PressFreedom <- PressFreedom |> pivot_longer(!(LOCATION | Country | Indicator)) |> rename("Time"="name", "Value"="value")
```





We now are cleaning the Religion variable.

First, we want to have a look at the columns names of our dataset.
```{r}
religion |> 
  names() |> 
  knitr::kable()
```

The Reliability variable only takes one value. 
```{r, warning=FALSE}
religion |> 
  distinct(Reliability) |> 
  knitr::kable()
```

Since our analysis is not based on gender disparity, we are going to only select "Both Sexes" values. Then, we will remove the columns we deem unnecessary and rename the remaining columns. 
```{r}
religion <- religion |> 
  filter(Sex=="Both Sexes") |> 
  select(-Sex, -Reliability, - "Value Footnotes", -"Source Year") |>  
  rename("Country" = "Country or Area", "Time" = "Year")
```





To finish, we now are cleaning the Institution variable, which only consists of renaming its columns. 
```{r}
institution <- institution |> 
  rename("Country" = "Entity", "LOCATION"= "Code", "Value"= "stability_dem_inst_bti")|>
  mutate("Indicator"= "Stability of Institutions")
```

## Data description and Tables 

The government expenditure on education is expressed in millions US$. 

This table has `r educ_expenditure |> ncol()` columns, which are `r educ_expenditure |> names()`.
It has `r educ_expenditure |> nrow()` rows and each of them represents a particular country at a particular year. Here is a table with basic statistics of the Expenditure on education variable.
```{r}
educ_expenditure |> 
  filter(Country != "Holy See") |> 
  group_by(Country) |> 
  summarise("min" = min(Value, na.rm=TRUE), "median" = median(Value, na.rm=TRUE), "max"= max(Value, na.rm=TRUE), "mean"= mean(Value, na.rm=TRUE)) |> 
  head(n=10) |> 
  knitr::kable(caption= "Statistics of expenditure on education per year")
```



The GDP and GDP per capita are expressed in terms of current US$. 

This GDP table has `r GDP |> ncol()` columns, which are `r GDP |> names()`. It has `r GDP |> nrow()` rows and each of them represents a particular country at a particular year. Here is a table with basic statistics of the GDP variable.  
```{r}
GDP |> 
  group_by(Country) |> 
  summarise("min" = min(Value, na.rm=TRUE), "median" = median(Value, na.rm=TRUE), "max"= max(Value, na.rm=TRUE), "mean"= mean(Value, na.rm=TRUE)) |> 
  head(n=10) |> 
  knitr::kable(caption= "Statistics of GDP")
```

This GDP per capita  table has `r GDP_capita |> ncol()` columns, which are `r GDP_capita |> names()`. It has `r GDP_capita |> nrow()` rows and each of them represents a particular country at a particular year. Here is a table with basic statistics of the GDP per capita variable. 
```{r}
GDP_capita |> 
  group_by(Country) |> 
  summarise("min" = min(Value, na.rm=TRUE), "median" = median(Value, na.rm=TRUE), "max"= max(Value, na.rm=TRUE), "mean"= mean(Value, na.rm=TRUE)) |> 
  head(n=10) |>
  knitr::kable(caption= "Statistics of GDP per capita per year")
```



Life expectancy at birth is expressed in total years.

This Life expectancy at birth table has `r life_expectancy|> ncol()` columns, which are `r life_expectancy |> names()`. It has `r life_expectancy|> nrow()` rows and each of them represents a particular country at a particular year. Here is a table with basic statistics of the Life expectancy at birth variable.  
```{r}
life_expectancy |> 
  group_by(Country) |> 
  summarise("min" = min(Value, na.rm=TRUE), "median" = median(Value, na.rm=TRUE), "max"= max(Value, na.rm=TRUE), "mean"= mean(Value, na.rm=TRUE)) |> 
  head(n=10) |> 
  knitr::kable(caption= "Statistics of life expectancy at birth (years) per year")
```



The population is expressed in thousands.

This total population table has `r population |> ncol()` columns, which are `r population |> names()`. It has `r population|> nrow()` rows and each of them represents a particular country at a particular year. Here is a table with basic statistics of the total population variable. 
```{r}
population |> 
  group_by(Country) |> 
  summarise("min" = min(Value, na.rm=TRUE), "median" = median(Value, na.rm=TRUE), "max"= max(Value, na.rm=TRUE), "mean"= mean(Value, na.rm=TRUE)) |> 
  head(n=10) |> 
  knitr::kable(caption= "Statistics of population per year")
```



The total net enrolment rate (in upper secondary, lower secondary and primary) of both sexes is expressed in percentage. 

This  total net enrolment rate table has `r enrolment |> ncol()` columns, which are `r enrolment |> names()`. It has `r enrolment |> nrow()` rows and each of them represents a particular country at a particular year. Here is a table with basic statistics of the total net enrolment rate at birth variable.  
```{r}
enrolment |> 
  filter(Country != "Holy See") |> 
  group_by(Country) |> 
  summarise("min" = min(Value, na.rm=TRUE), "median" = median(Value, na.rm=TRUE), "max"= max(Value, na.rm=TRUE), "mean"= mean(Value, na.rm=TRUE)) |> 
  head(n=10) |> 
  knitr::kable(caption= "Statistics of enrolment per year")
```



The Rural population is expressed as a percentage of the total population.

This Rural population table has `r rural |> ncol()` columns, which are `r rural |> names()`. It has `r rural |> nrow()` rows and each of them represents a particular country at a particular year. Here is a table with basic statistics of the Rural population variable.
```{r, warning= FALSE}
rural |> 
  group_by(Country) |> 
  summarise("min" = min(Value, na.rm=TRUE), "median" = median(Value, na.rm=TRUE), "max"= max(Value, na.rm=TRUE), "mean"= mean(Value, na.rm=TRUE)) |> 
  head(n=10) |> 
  knitr::kable(caption= "Statistics of rural population share per year")
```



The access to electricity is measured as a percentage of the population.

This access to electricity table has `r electricity |> ncol()` columns, which are `r electricity |> names()`. It has `r electricity |> nrow()` rows and each of them represents a particular country at a particular year. Here is a table with basic statistics of the access to electricity variable.  
```{r}
electricity |> 
  group_by(Country) |> 
  filter(!(Country %in% c("Kosovo", "American Samoa", "Not classified"))) |>
  summarise("min" = min(Value, na.rm=TRUE), "median" = median(Value, na.rm=TRUE), "max"= max(Value, na.rm=TRUE), "mean"= mean(Value, na.rm=TRUE)) |> 
  head(n=10) |> 
  knitr::kable(caption= "Statistics of electricity access per year")

```



The press freedom table shows for each country their rank and index. 

This  press freedom table has `r PressFreedom |> ncol()` columns, which are `r PressFreedom |> names()`. It has `r PressFreedom |> nrow()` rows and each of them represents a particular country at a particular year. Here is a table with basic statistics of the  press freedom variable.  
```{r}
PressFreedom |> group_by(Country) |> 
  summarise("min" = min(Value, na.rm=TRUE), "median" = median(Value, na.rm=TRUE), "max"= max(Value, na.rm=TRUE), "mean"= mean(Value, na.rm=TRUE)) |> 
  head(n=10) |> 
  knitr::kable(caption= "Statistics of freedom of press per year")
```



The stability of institutions of each country is represented with the Bertelsmann Transformation Index (BTI). This index is a global ranking and analysis of political and economic transformation in developing countries. It is published by the Bertelsmann Stiftung, a German private foundation.

This stability of institutions table has `r institution |> ncol()` columns, which are `r institution |> names()`. It has `r institution|> nrow()` rows and each of them represents a particular country at a particular year. Here is a table with basic statistics of the stability of institutions variable. 
```{r}
institution |> group_by(Country) |> 
  summarise("min" = min(Value, na.rm=TRUE), "median" = median(Value, na.rm=TRUE), "max"= max(Value, na.rm=TRUE), "mean"= mean(Value, na.rm=TRUE)) |> 
  head(n=10) |> 
  knitr::kable(caption= "Statistics of institution stability per year")
```



The religion table shows the exact number of believers for every religion in each countries display per sexes and as a total (i.e both sexes).

This Religion table has `r religion |> ncol()` columns, which are `r religion |> names()`. It has `r religion |> nrow()` rows and each of them represents in each country the number of a certain religion in a particular year. Here is a table which gives the most practised religion by country and year.
```{r}
religion |> filter(Religion != "Total") |> 
  group_by(Country, Time) |> 
  slice_max(order_by = Value) |> 
  select(Religion, Country, Time) |> 
  head(n=10) |> 
  knitr::kable(caption = "Most popular religion by country and year")
```

## Research question description 

Nowadays, the majority of individuals attain a secondary education, marking a notable shift from historical trends. Illiteracy, defined as the incapacity to read or write, was widespread in 1820, with only 12% of the global population possessing these skills. Since then, a significant reversal has occurred. Over the past 65 years, the global literacy rate has consistently increased by 4% every five years, rising from 42% in 1960 to 86% in 2015. Despite this progress, a considerable segment of the population remains illiterate, constituting 14% of the world population as of 2016.The causes behind this persistent illiteracy are subject to debate, with factors such as poverty, population growth, and ineffective government policies often cited. 

However, in our study, we refrain from addressing the "why" question. The scale and methodologies employed are not sufficiently robust for us to draw causal conclusions. Instead, our approach is more observational, focusing on the "who" aspect. Our objective is to identify common characteristics among illiterate individuals, allowing us to observe patterns in this demographic.  To facilitate our study, we opt to concentrate on a specific subset of the illiterate population, namely, the young illiterates (defined here as 15-24 years olds). This demographic will be analysed on a country-by-country basis. Essentially, our aim is to assert whether there are commonalities in countries experiencing high levels of youth illiteracy. To achieve this, we have selected multiple variables that we believe provide insightful information, including institutional stability, overall population, rural population, GDP, GDP per capita, access to electricity, religion, and the percentage of enrolment by educational level, among others.  

At the end of our work, we anticipate two possible outcomes. First, no discernible patterns emerge, which will tell us that the distribution of the illiterate population is genuinely random and relatively evenly spread across the global population. Second, similarities are observed among countries with high levels of illiteracy which will allow us to speculate on potential explanations for illiteracy.

## Graphical representation of our target variable

First, we want to have a look at the distribution oveer time of Illiterate people (15-24 years old) around the world.
```{r}
illiteracy |> filter(Country=="World") |> ggplot(aes(x= Time, y=Value/10^3)) + geom_col() + ylab("Number of illiterate people around the world") + ggtitle("Distribution of Illiteracy (in thousands) over time")
```
